Denotational Semantics of General Payment Primitives, and Its Payment System

### 2022-09-22 packages/spec-haskell/yellowpaper/Paper.tex
DONE subsection: Money Distribution Models

But when it comes to the function of money, it is less controversial. According to Von Mises,

@@

This yellow paper adds a new controversy to money, that is to present a survey challenging the preconceived notion of
how money can perform its function of medium of exchange using computer science.

In Chapter 2, we shall first explore the foundation for the survey. Here we present a formal definition of payment system,
and various models used in the system. We then select a few relevant approaches used in computer science useful for
modeling and defining formal specification for the payment system.

@@

We first define a payment system, whereby:

@@

@@

\begin{itemize}
\item U is the set of monetary units
\item $\nu : U \rightarrow \mathbb{N}$ is the value function defining the value $\nu(u)$ of every value unit u. The set
    $\mathbb{N}$ is the set of all natural numbers, but instead, we can use any set of numerals that is totally ordered
    (e.g.\ integers, real numbers).
\item $\beta : U \rightarrow \mathfrak{B}$ is the bearer function defining the bearer $\beta(u)$ of a unit. The set
    $\mathfrak{B}$ is the set of possible bearers. The bearer is usually a legal construction defining any type of legal
    entity, such as a person, a family, a company, a state institution, etc.
\end{itemize}

This discrete nature of money distribution is schematically depicted in {\ref{fig:discrete-md}}:

@@

\subsubsection{A New Unifying Model}

@@

>>>>

This yellow paper does not intend to address these controversies, but to focus on the function of money. According to
Von Mises:

@@

This paper adds a new controversy to money, that is to present a survey challenging the preconceived notion of how money
can perform its function of medium of exchange using computer science.

In Chapter 2, we shall first explore the foundation for the survey. Here we present a formal definition of payment
system and its components. We then select a few relevant approaches used in computer science useful for modeling and
defining formal specification for the payment system.

@@

Here we present a formal definition of payment system and its components.

\paragraph{}

A payment system can be solely defined by these components:

@@

\paragraph{}

@@

\begin{itemize}
\item U is the set of monetary units.
\item $\nu : U \rightarrow \mathbb{N}$ is the value function defining the value $\nu(u)$ of every value unit u. The set
    $\mathbb{N}$ is the set of all natural numbers, but instead, we can use any set of numerals that is totally ordered
    (e.g.\ integers, real numbers).
\item $\beta : U \rightarrow \mathfrak{B}$ is the bearer function defining the bearer $\beta(u)$ of a unit. The set
    $\mathfrak{B}$ is the set of possible bearers. The bearer is usually a legal construction defining any type of legal
    entity, such as a person, a family, a company, a state institution, etc.
\end{itemize}

This discrete nature of this money distribution model is schematically depicted in {\ref{fig:discrete-md}}:

@@

\subsubsection{Using Shared Context}

But the discrete nature of the model does not provide the necessary element for us to add the desired properties to the
payment system. Here we propose a modification that involves the usage of \textit{shared context (ctx for short)}:

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{../assets/money-distribution-with-ctx.png}
    \caption{Schematic representation of money distribution with shared context}
    \label{fig:md-with-ctx}
\end{figure}

Note that in this model, shared context can be updated independently while value functions of the same money
distribution can produce different monetary values.

\paragraph{}

Here are some examples how that could work:

\begin{itemize}
\item If time is included the shared context, then monetary value of each money unit can be continuous over time.
\item A subset of money units can also have their value functions depending on the same information in shared
    context. This could enable payment primitives that involve many parties.
\end{itemize}

\subsubsection{Haskell Definitions}

@@

An attempt of creating a unifying theory for money and its distribution was made by Buldas, Saarepera et al.\cite{buldas2021unifying}

\begin{displayquote}
A useful observation about existing money schemes is that they all have some kind of monetary units that are physical or
digital representations of money. Examples are bills, coins, bank accounts, Bitcoin UTXOs, etc.
\end{displayquote}
